Step1_LoadData

### 2024-07-16 ProxySelection.Rmd
html
prxy

---
title: "Step1_LoadData"
author: "Yang Lei"
date: "2024-07-11"
output: pdf_document
---

@@

**Method 1: Bross Formula**
```{r, warning=FALSE, cache=TRUE}
out3 <- get_prioritised_covariates(df = out2,
                                   patientIdVarname = "idx", 
                                   exposureVector = basetable$exposure,
                                   outcomeVector = basetable$outcome,
                                   patientIdVector = patientIds, 
                                   k = 100)
hdps.dim <- out3$autoselected_covariate_df

@@

print("The Odds Ratio by Bross Formula")
round(sum_ci.bross, 3)
```

**Method 2: Hybrid Bross Formula & Lasso**
```{r, warning=FALSE, cache=TRUE}
set.seed(42)
proxy.list <- names(out3$autoselected_covariate_df[,-1])
covarsTfull <- c(investigator.specified.covariates, proxy.list)
Y.form <- as.formula(paste0(c("outcome~ exposure", 
                              covarsTfull), collapse = "+") )
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]
lasso.fit<-glmnet::cv.glmnet(y = hdps.data$outcome, 
                             x = covar.mat, 
                             type.measure='mse',
                             family="binomial",
                             alpha = 1, 
                             nfolds = 5)
coef.fit<-coef(lasso.fit,s='lambda.min',exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.hybrid <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.hybrid)
fs.hybrid <- proxy.list.sel.hybrid

@@

print("The Odds Ratio by Bross Formula & Lasso")
round(sum_ci.hybrid, 3)
```

_Before implementing other methods, drop the information from method 1 (Bross formula). Continue with the results from Step 3: Recurrence (table `out2`)._
```{r, warning=FALSE, cache=TRUE}
hdps.dim <- out2

@@

**Method 3: Pure Lasso**
```{r, warning=FALSE, cache=TRUE}
set.seed(42) # different seed -> different length for the selected proxy list
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]
lasso.fit <- glmnet::cv.glmnet(y = hdps.data$outcome, 
                               x = covar.mat, 
                               type.measure='mse',
                               family="binomial",
                               alpha = 1, 
                               nfolds = 5)
coef.fit<-coef(lasso.fit,s='lambda.min',exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.ml <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.ml)
fs.lasso <- proxy.list.sel.ml
proxy.lasso <- paste0(fs.lasso, collapse = "+")
form.lasso <- paste0(c(covform, proxy.lasso), collapse = "+")
formula.lasso <- as.formula(paste0("exposure", "~", form.lasso))

@@

# Q2: Odds ratio
print("The Odds Ratio by Method Pure Lasso")
sum_ci.lasso
```

**Method 4: Elastic Net**
```{r, warning=FALSE, cache=TRUE}
set.seed(42)
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]
elas.fit <- glmnet::cv.glmnet(y = hdps.data$outcome, 
                             x = covar.mat, 
                             type.measure='mse',
                             family="binomial",
                             alpha = 0.5, 
                             nfolds = 5)
coef.fit<-coef(elas.fit, s='lambda.min', exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.elastic <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.elastic)
fs.elastic <- proxy.list.sel.elastic

@@

# Q2: Odds ratio
print("The Odds Ratio by Elastic Net")
sum_ci.elastic
```

**Method 5: XGBoost**
```{r, warning=FALSE, cache=TRUE}
xgb_fit <- xgboost(data = covar.mat, label = hdps.data$outcome, 
                    max.depth = 20, eta = 1, nthread = 2, nrounds = 5, 
                    objective = "binary:logistic")

@@

# Q2: Odds ratio
print("The Odds Ratio by XGBoost")
sum_ci.xgboost
```

@@

_GA (Genetic Algorithm) has no specific internal model building mechanism and would wrap around whatever model the user specifies_
```{r, warning=FALSE, cache=TRUE}
gaOpt <- function(vars, IV.train, DV.train) {
	varNames <- colnames(IV.train) #getting names of all variables
	selectedVarNames <- varNames[vars == "1"] # getting names of selected vars from GA
	gaSolutionData <- IV.train[,selectedVarNames] # keeping only those selected vars

@@

# Q2: Odds ratio
print("The Odds Ratio by GA:")
sum_ci.GA
```

@@

_(Long Running Time)_
```{r, warning=FALSE, cache=TRUE}
svmTrainOutcome <- hdps.data$outcome
svmTrainOutcome[svmTrainOutcome == 0] <- -1
svmTrainPreds <- data.frame(covar.mat)

@@

# Q2: Odds ratio
print("The Odds Ratio by Elastic SCAD SVM")
sum_ci.svm
```

@@

_(Long Running Time)_
```{r, warning=FALSE, cache=TRUE}
set.seed(42)
boruta_output <- Boruta(Y.form, data = hdps.data, doTrace = 1)
sel.variables.bor <- names(boruta_output$finalDecision)[which(boruta_output$finalDecision != "Rejected")]
fs.boruta <- proxy.list[proxy.list %in% sel.variables.bor]
proxy.boruta <- paste0(fs.boruta, collapse = "+")
form.boruta <- paste0(c(covform, proxy.boruta), collapse = "+")
formula.boruta <- as.formula(paste0("exposure", "~", form.boruta))

@@

# Q2: Odds ratio
print("The Odds Ratio by Boruta Selection")
sum_ci.boruta
```

**Method 9.1: Stepwise Regression (Forward)**
```{r, warning=FALSE, cache=TRUE}
set.seed(42)

initial_formula <- as.formula(paste0("outcome~exposure+",
                                     covform,
                                     collapse = "+"))
full_formula <- as.formula(paste0("outcome~exposure+",
                                  rhsformula,
                                  collapse = "+"))

@@

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Forward Selection")
sum_ci.forward
```

**Method 9.2: Stepwise Regression (Backward)**
```{r, warning=FALSE, cache=TRUE}
stepwise_backward <- stepAIC(full_model, direction = "backward")
fs.backward <- proxy.list[proxy.list %in% fs.backward]
fs.backward <- fs.backward[fs.backward != "exposure"]
proxy.backward <- paste0(fs.backward, collapse = "+")
form.backward <- paste0(c(covform, proxy.backward), collapse = "+")
formula.backward <- as.formula(paste0("exposure", "~", form.backward))

@@

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Backward Selection")
sum_ci.backward
```

**Method 9.3: Stepwise Regression (Both)**
```{r, warning=FALSE, cache=TRUE}
stepwise_both <- stepAIC(initial_model, 
                            scope = list(lower = initial_model, 
                                         upper = full_model), 
                            direction = "both")
fs.both <- all.vars(formula(stepwise_both))[-1]
fs.both <- proxy.list[proxy.list %in% fs.both]
proxy.both <- paste0(fs.both, collapse = "+")
form.both <- paste0(c(covform, proxy.both), collapse = "+")
formula.both <- as.formula(paste0("exposure", "~", form.both))

@@

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Both Selection")
sum_ci.both
```

Comparing the Results:
```{r, warning=FALSE, cache=TRUE}
sum_ci.all <- rbind(sum_ci.bross, sum_ci.hybrid, sum_ci.lasso, sum_ci.elastic, 
                    sum_ci.xgboost, sum_ci.GA, sum_ci.svm, sum_ci.boruta,
                    sum_ci.forward, sum_ci.backward, sum_ci.both)

@@

Print the comparing plot & table:
```{r}
print("Comparison of odds ratios and the confidence intervals by different proxy selection methods:")
OR.ci_plot
```

>>>>

---
title: "Step1_LoadData"
author: "Yang Lei"
date: "2024-07-11"
output:
  html_document:
    df_print: paged
---

@@

**Method 1: Bross Formula**
```{r bross, warning=FALSE, cache=TRUE}
start_time <- Sys.time()
out3 <- get_prioritised_covariates(df = out2,
                                   patientIdVarname = "idx", 
                                   exposureVector = basetable$exposure,
                                   outcomeVector = basetable$outcome,
                                   patientIdVector = patientIds, 
                                   k = 100)
hdps.dim <- out3$autoselected_covariate_df

@@

print("The Odds Ratio by Bross Formula")
round(sum_ci.bross, 3)
end_time <- Sys.time()
elapsed_time.bross <- as.numeric(difftime(end_time, start_time, units = "secs"))
```

**Method 2: Hybrid Bross Formula & Lasso**
```{r hybrid, warning=FALSE, cache=TRUE}
set.seed(42)
start_time <- Sys.time()
proxy.list <- names(out3$autoselected_covariate_df[,-1])
covarsTfull <- c(investigator.specified.covariates, proxy.list)
Y.form <- as.formula(paste0(c("outcome~ exposure", 
                              covarsTfull), collapse = "+") )
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]
lasso.fit<-glmnet::cv.glmnet(y = hdps.data$outcome, 
                             x = covar.mat, 
                             type.measure='mse',
                             family="binomial",
                             alpha = 1, 
                             nfolds = 5)
coef.fit<-coef(lasso.fit,s='lambda.min',exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.hybrid <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.hybrid)
fs.hybrid <- proxy.list.sel.hybrid

@@

print("The Odds Ratio by Bross Formula & Lasso")
round(sum_ci.hybrid, 3)
end_time <- Sys.time()
elapsed_time.hybrid <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.hybrid
```

_Before implementing other methods, drop the information from method 1 (Bross formula). Continue with the results from Step 3: Recurrence (table `out2`)._
```{r prepare, warning=FALSE, cache=TRUE}
hdps.dim <- out2

@@

**Method 3: Pure Lasso**
```{r lasso, warning=FALSE, cache=TRUE}
set.seed(42) # different seed -> different length for the selected proxy list
start_time <- Sys.time()
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]
lasso.fit <- glmnet::cv.glmnet(y = hdps.data$outcome, 
                               x = covar.mat, 
                               type.measure='mse',
                               family="binomial",
                               alpha = 1, 
                               nfolds = 5)
coef.fit<-coef(lasso.fit,s='lambda.min',exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.ml <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.ml)
fs.lasso <- proxy.list.sel.ml
proxy.lasso <- paste0(fs.lasso, collapse = "+")
form.lasso <- paste0(c(covform, proxy.lasso), collapse = "+")
formula.lasso <- as.formula(paste0("exposure", "~", form.lasso))

@@

# Q2: Odds ratio
print("The Odds Ratio by Method Pure Lasso")
sum_ci.lasso
end_time <- Sys.time()
elapsed_time.lasso <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.lasso
```

**Method 4: Elastic Net**
```{r enet, warning=FALSE, cache=TRUE}
set.seed(42)
start_time <- Sys.time()
covar.mat <- model.matrix(Y.form, data = hdps.data)[,-1]
elas.fit <- glmnet::cv.glmnet(y = hdps.data$outcome, 
                             x = covar.mat, 
                             type.measure='mse',
                             family="binomial",
                             alpha = 0.5, 
                             nfolds = 5)
coef.fit<-coef(elas.fit, s='lambda.min', exact=TRUE)
sel.variables<-row.names(coef.fit)[which(as.numeric(coef.fit)!=0)]
proxy.list.sel.elastic <- proxy.list[proxy.list %in% sel.variables]
length(proxy.list.sel.elastic)
fs.elastic <- proxy.list.sel.elastic

@@

# Q2: Odds ratio
print("The Odds Ratio by Elastic Net")
sum_ci.elastic
end_time <- Sys.time()
elapsed_time.enet <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.enet
```

**Method 5: XGBoost**
```{r xgb, warning=FALSE, cache=TRUE}
start_time <- Sys.time()
xgb_fit <- xgboost(data = covar.mat, label = hdps.data$outcome, 
                    max.depth = 20, eta = 1, nthread = 2, nrounds = 5, 
                    objective = "binary:logistic")

@@

# Q2: Odds ratio
print("The Odds Ratio by XGBoost")
sum_ci.xgboost
end_time <- Sys.time()
elapsed_time.xgb <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.xgb
```

@@

_GA (Genetic Algorithm) has no specific internal model building mechanism and would wrap around whatever model the user specifies_
```{r ga, warning=FALSE, cache=TRUE}
start_time <- Sys.time()
gaOpt <- function(vars, IV.train, DV.train) {
	varNames <- colnames(IV.train) #getting names of all variables
	selectedVarNames <- varNames[vars == "1"] # getting names of selected vars from GA
	gaSolutionData <- IV.train[,selectedVarNames] # keeping only those selected vars

@@

# Q2: Odds ratio
print("The Odds Ratio by GA:")
sum_ci.GA
end_time <- Sys.time()
elapsed_time.ga <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.ga
```

@@

_(Long Running Time)_
```{r svm, warning=FALSE, cache=TRUE}
start_time <- Sys.time()
svmTrainOutcome <- hdps.data$outcome
svmTrainOutcome[svmTrainOutcome == 0] <- -1
svmTrainPreds <- data.frame(covar.mat)

@@

# Q2: Odds ratio
print("The Odds Ratio by Elastic SCAD SVM")
sum_ci.svm
end_time <- Sys.time()
elapsed_time.svm <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.svm
```

@@

_(Long Running Time)_
```{r Boruta, warning=FALSE, cache=TRUE}
set.seed(42)
start_time <- Sys.time()
boruta_output <- Boruta(Y.form, data = hdps.data, doTrace = 1)
sel.variables.bor <- names(boruta_output$finalDecision)[which(boruta_output$finalDecision != "Rejected")]
fs.boruta <- proxy.list[proxy.list %in% sel.variables.bor]
proxy.boruta <- paste0(fs.boruta, collapse = "+")
form.boruta <- paste0(c(covform, proxy.boruta), collapse = "+")
formula.boruta <- as.formula(paste0("exposure", "~", form.boruta))

@@

# Q2: Odds ratio
print("The Odds Ratio by Boruta Selection")
sum_ci.boruta
end_time <- Sys.time()
elapsed_time.boruta <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.boruta
```

**Method 9.1: Stepwise Regression (Forward)**
```{r fs, warning=FALSE, cache=TRUE}
set.seed(42)
start_time <- Sys.time()
initial_formula <- as.formula(paste0("outcome~exposure+",
                                     covform,
                                     collapse = "+"))
full_formula <- as.formula(paste0("outcome~exposure+",
                                  rhsformula,
                                  collapse = "+"))

@@

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Forward Selection")
sum_ci.forward
end_time <- Sys.time()
elapsed_time.fs <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.fs
```

**Method 9.2: Stepwise Regression (Backward)**
```{r bs, warning=FALSE, cache=TRUE}
start_time <- Sys.time()
stepwise_backward <- stepAIC(full_model, direction = "backward")
fs.backward <- proxy.list[proxy.list %in% fs.backward]
fs.backward <- fs.backward[fs.backward != "exposure"]
proxy.backward <- paste0(fs.backward, collapse = "+")
form.backward <- paste0(c(covform, proxy.backward), collapse = "+")
formula.backward <- as.formula(paste0("exposure", "~", form.backward))

@@

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Backward Selection")
sum_ci.backward
end_time <- Sys.time()
elapsed_time.bs <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.bs
```

**Method 9.3: Stepwise Regression (Both)**
```{r sw, warning=FALSE, cache=TRUE}
start_time <- Sys.time()
stepwise_both <- stepAIC(initial_model, 
                            scope = list(lower = initial_model, 
                                         upper = full_model), 
                            direction = "both")
fs.both <- all.vars(formula(stepwise_both))[-1]
fs.both <- proxy.list[proxy.list %in% fs.both]
proxy.both <- paste0(fs.both, collapse = "+")
form.both <- paste0(c(covform, proxy.both), collapse = "+")
formula.both <- as.formula(paste0("exposure", "~", form.both))

@@

# Q2: Odds ratio
print("The Odds Ratio by Stepwise Both Selection")
sum_ci.both
end_time <- Sys.time()
elapsed_time.sw <- as.numeric(difftime(end_time, start_time, units = "secs"))
elapsed_time.sw
```

Comparing the Results:
```{r compare, warning=FALSE, cache=TRUE}
sum_ci.all <- rbind(sum_ci.bross, sum_ci.hybrid, sum_ci.lasso, sum_ci.elastic, 
                    sum_ci.xgboost, sum_ci.GA, sum_ci.svm, sum_ci.boruta,
                    sum_ci.forward, sum_ci.backward, sum_ci.both)

@@

Print the comparing plot & table:
```{r}
elapsed_times <- list()
elapsed_times[["bross"]] <- elapsed_time.bross
elapsed_times[["hybrid"]] <- elapsed_time.hybrid
elapsed_times[["lasso"]] <- elapsed_time.lasso
elapsed_times[["enet"]] <- elapsed_time.enet
elapsed_times[["xgb"]] <- elapsed_time.xgb
elapsed_times[["ga"]] <- elapsed_time.ga
elapsed_times[["svm"]] <- elapsed_time.svm
elapsed_times[["boruta"]] <- elapsed_time.boruta
elapsed_times[["fs"]] <- elapsed_time.fs
elapsed_times[["bs"]] <- elapsed_time.bs
elapsed_times[["sw"]] <- elapsed_time.sw
print(elapsed_times)
print("Comparison of odds ratios and the confidence intervals by different proxy selection methods:")
OR.ci_plot
```
